An Overview on Temporal Databases.\\ New Approaches and Proposals.

### 2012-03-01 book/sources/intro.tex
Almost done!
The new form of the chapter...

The concept of time itself is very complex to handle and interpret~\cite{klein94},~\cite{Shackle61}, though it is very natural and omnipresent. As information systems often attempt the modelling or the representation of natural objects, concepts or processes, they often require modelling or representing temporal information. Thus, several proposals have been concerned with the obtaining of theoretical models that allow the modelling or representation of time~\cite{Bolour82},~\cite{Cru97}.

A very specific type of information systems are database systems, which are computer systems designed to manage databases. A database contains data representing real objects or concepts. Every one of these data is a measurement or description of a property of a real object or concept. In reality, some aspects or properties of objects or concepts are time-variant or time-related. E.g. the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant. A temporal database schema is a database schema that models real objects or concepts with time-related or time-variant properties. However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints. An example. Consider a relation in a relational library database, modelling the presence of books in the library. Every physical book is represented by a unique identifier. Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it was returned). Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates. This would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible. A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.

A lot of research concerns these temporal database models and their approaches to the modelling of time. The first efforts were towards the representation of historical information related to objects represented by records in a database~\cite{Clifford:1985:AHR:971699.318922}. Some proposals tried to extend the Entity Relationship Model~\cite{Klopprogge:1983}, without impact on any database standards like SQL~\cite{Sarda:1990:ESH:627277.627409}.

@@

Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}. Humans are often not hindered by imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions. Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling. E.g. the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 1/01/1201 A.D. and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression. It is known that the building was finished on a single day, but it is not known precisely which day this was.

@@

The rest of this chapter is structured as follows. Section \ref{sec:time-domain} presents some basic concepts and terminology about temporal modelling and discusses some of its important aspects and issues. In section \ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems. \textcolor{red}{TODO: REFERENCE SECTION 4!}

>>>>

The concept of time itself is very complex to handle and interpret~\cite{klein94},~\cite{Shackle61}, though it is very natural and omnipresent. As information systems often attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts. Thus, several proposals have been concerned with the obtaining of theoretical models that allow the modelling or representation of time~\cite{Bolour82},~\cite{Cru97}.

A very specific type of information systems are database systems, which are computer systems designed to manage databases. A database contains data representing real objects or concepts. Every one of these data is a result value of a measurement of a property or a description of a property of a real object or concept. In reality, some aspects or properties of objects or concepts are time-variant or time-related. E.g. the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant. A temporal database schema is a database schema that models real objects or concepts with time-related or time-variant properties. However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints. An example. Consider a relation in a relational library database, modelling the presence of books in the library. Every physical book is represented by a unique identifier. Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it was returned). Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates. This group of records would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible. A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.

A lot of research concerns temporal database models and their approaches to the modelling of time. The first efforts were towards the representation of historical information related to objects represented by records in a database~\cite{Clifford:1985:AHR:971699.318922}. Some proposals tried to extend the Entity Relationship Model~\cite{Klopprogge:1983}, without impact on any database standards like SQL~\cite{Sarda:1990:ESH:627277.627409}.

@@

Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they are often not hindered by imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions. Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling. E.g. the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 1/01/1201 A.D. and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression. It is known that the building was finished on a single day, but it is not known precisely which day this was.

@@

The rest of this chapter is structured as follows. Section \ref{sec:time-domain} presents some basic concepts and terminology about temporal modelling and discusses some of its important aspects and issues. In section \ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems. Finally, in section \ref{sec:imprecision-uncertainty}, an approach to querying temporal databases containing imperfect temporal information is presented, followed by some conclusions and some suggestions for future work in section \ref{sec:further-research-work}.
